### 2022-09-27 main/chapter_context.tex
fiddle with cosmology a little; add sakharov

Early versions of the Standard Model had only two fermion generations and
no neutrino masses~\cite{
wells2020discovery,
bjorken1985november
},
but the third generation and non-zero neutrino masses are now observed;
for example, the third generation is shown by $\mathrm{CP}$ asymmetry in
kaons~\cite{
cronin1964evidence,
kabayasji1973cpv
},
precise measurements of the $Z$ boson~\cite{
lep2006precision
},
and direct observations~\cite{
perl1977evidence,
herb1977observation,
abachi1995observation
};
neutrino masses are also necessary to explain their observed flavour
oscillations~\cite{
kamiokande1998measurement,
superk1998evidence,
superk1999measurement,
lsnd1998evidence
}.

@@

When `nobody knows' the answer to a riddle, the status lies between two
extremes:
many answers appear satisfactory and none is the clear best,
or no satisfactory answer is yet known.
We distinguish these two cases with the respective labels
\emph{uncertain} and \emph{unknown}:
the model of neutrino masses is mostly \emph{uncertain}, but
the reconciliation quantum field theory with general relativity is
mostly \emph{unknown}.

@@

\TODO{
Three more mysteries:
cosmological constant~\cite{riess1998observational},
inflation~\cite{bezrukov2008higgs},
baryogenesis (matter-antimatter asymmetry)
}

\paragraph{Origins.}

>>>>

Early versions of the Standard Model had only two fermion generations and
no neutrino masses~\cite{
wells2020discovery,
bjorken1985november
},
but the third generation and non-zero neutrino masses are now observed;
for example, the third generation is shown by charge-parity $CP$ asymmetry in
kaons~\cite{
cronin1964evidence,
kabayasji1973cpv
},
precise measurements of the $Z$ boson~\cite{
lep2006precision
},
and direct observations~\cite{
perl1977evidence,
herb1977observation,
abachi1995observation
};
neutrino masses are also necessary to explain their observed flavour
oscillations~\cite{
kamiokande1998measurement,
superk1998evidence,
superk1999measurement,
lsnd1998evidence
}.

@@

When `nobody knows' the answer to a riddle, the status lies somewhere between
two extremes:
either many answers appear satisfactory and none is the clear best,
or no satisfactory answer is yet known.
We distinguish these two cases with the respective labels
\emph{uncertain} and \emph{unknown}:
the model of neutrino masses is mostly \emph{uncertain}, but
the reconciliation quantum field theory with general relativity is
mostly \emph{unknown}.

@@

Other mysteries of cosmology include:
the positive cosmological constant,
the inflation of the early universe,
and baryogenesis~\cite{
wells2020discovery,
riess1998observational
}.
Each lies somewhere between uncertain and unknown, but may not be entirely
out of reach;
the Higgs field could plausibly have driven
inflation~\cite{bezrukov2008higgs},
and the Standard Model does slightly violate $C$ and $CP$, both of which
are required for baryogenesis~\cite{sakharov1991re}.

\paragraph{Origins.}
\TODO{parameters and structure}
